---
title: "FMPH-400_Biostats_Final"
output:
  pdf_document: default
  word_document: default
---

## Set Up ##

```{r setup, include=FALSE}
#Working Directory
getwd()
setwd("/Users/rj/Documents/School/UCSD/Fall_2021/FMPH-400_Biostatistics")

knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8, fig.path='Fig/')

#Read in data
library(haven)

Hypertens <- read_sav("Hypertension.sav")

```

## Data Analysis ##

#1. Association between diabetes and systolic blood pressure

```{r}
## Diabetes and SBP.Change ##

# Subset
DiabetesCH <- Hypertens$SBP.Change[Hypertens$Diabetes == 1]
NoDiabetesCH <- Hypertens$SBP.Change[Hypertens$Diabetes == 0]

# Check Variance
sd(DiabetesCH)/sd(NoDiabetesCH) # = 0.994
sd(NoDiabetesCH)/sd(DiabetesCH) # = 1.006

# Five Number Summary
summary(DiabetesCH)
summary(NoDiabetesCH)

# Standard Deviation
sd(DiabetesCH)
sd(NoDiabetesCH)

# Counts for Individuals with Diabetes and no Diabetes
table(Hypertens$Diabetes)
```

```{r}
# Boxplot of systolic blood pressure change in diabetes group and non-diabetes group
boxplot(Hypertens$SBP.Change~Hypertens$Diabetes, names = c("Non-Diabetic", "Diabetic"), 
main="SBP Change in Diabetics vs. Non-Diabetics", 
xlab = "Presence or Absence of Diabetes", ylab = "Systolic Blood Pressure Change",
        col=c("DarkRed", "Orange"))
```


```{r}
# T-test for difference in means of Diabetes
t.test(NoDiabetesCH, DiabetesCH, var.equal = TRUE)
```

p-value = 0.200 --> not < 0.05, fail to reject null
95% CI = -7.46 to 1.56 --> contains 0, fail to reject null
t-stat = -1.28 --> not within rejection region of < -1.96 or  > 1.96, fail to reject null

#2. Association between homelessness and systolic blood pressure

```{r}
## Homeless and SBP.Change ##

# Subset
HomelessCH <- Hypertens$SBP.Change[Hypertens$Homeless == 1]
NoHomelessCH <- Hypertens$SBP.Change[Hypertens$Homeless == 0]

# Check Variance
sd(HomelessCH)/sd(NoHomelessCH) # = 0.989
sd(NoHomelessCH)/sd(HomelessCH) # = 1.010

# Summary
summary(HomelessCH)
summary(NoHomelessCH)

# Standard Deviation
sd(HomelessCH)
sd(NoHomelessCH)

# Counts for Individuals who are Homeless and not Homeless
table(Hypertens$Homeless)
```

```{r}
# Boxplot of systolic blood pressure change in Homeless group and non-Homeless groups
boxplot(Hypertens$SBP.Change~Hypertens$Homeless, names = c("Housed", "Homeless"), main = "SBP Change in Homeless vs. Non-Homeless", 
        xlab = "Housing-Type Groups", ylab = "Systolic Blood Pressure Change",
        col=c("DarkRed", "Orange"))
```


```{r}
# T-test for difference in means of Homeless
t.test(NoHomelessCH, HomelessCH, var.equal = TRUE)
```

p-value = 0.825 --> not < 0.05, fail to reject null
95% CI = -4.70 to 5.90 --> contains 0, fail to reject null
t-stat = 0.22 --> --> not within rejection region of < -1.96 or  > 1.96, fail to reject null

#3. Association between gender and systolic blood pressure

```{r}
## Gender and SBP.Change ##

# Subset
MaleCH <- Hypertens$SBP.Change[Hypertens$Gender == 0]
FemaleCH <- Hypertens$SBP.Change[Hypertens$Gender == 1]

# Check Variance
sd(MaleCH)/sd(FemaleCH) # = 0.929
sd(FemaleCH)/sd(MaleCH) # = 1.076 

# Summary
summary(MaleCH)
summary(FemaleCH)

# Standard Deviation
sd(MaleCH)
sd(FemaleCH)

# Count of Male and Female
table(Hypertens$Gender)

```

```{r}
# Boxplot of change in systolic blood pressure between Males and Females
boxplot(Hypertens$SBP.Change~Hypertens$Gender,
        names = c("Male", "Female"),
        main="SBP Change in Males vs. Females",
        xlab = "Gender Groups", ylab = "Systolic Blood Pressure Change", 
        col=c("DarkRed", "Orange"))
```


```{r}
# T-test for difference in means of Gender
t.test(MaleCH, FemaleCH, var.equal = TRUE)
```

p-value = 0.797 --> not < 0.05, fail to reject null
95% CI = -3.953 to 5.13 --> contains 0, fail to reject null
t-stat = 0.256 --> not within rejection region of < -1.96 or  > 1.96, fail to reject null


